Symmetric Peer-to-Peer Applications

### 2022-10-20 tpd-22.tex
...

As illustrated in Figure~\ref{fig.middleware}, events are broadcast between
peers with a timestamp scheduled to the future such that all peers are able to
apply them in sync.
%
To prevent broadcast cycles, each peer keeps a vector with the highest
timestamp it received from each other peer, such that lower numbers can be
ignored when received.
Since we rely on TCP connections, it is not possible to receive out-of-order
timestamps attached to a given source peer.
When the middleware receives a higher number from a given peer, it updates its
vector position and triggers a broadcast to neighbours.
%
As mentioned in Section~\ref{sec.tml.api}, we assume that peers assign
themselves unique identifiers in a contiguous range, such that they fit in a
contiguous vector of timestamps.
Therefore, this is not a severe limitation, and a separate mechanism could be
used for assignments, with few modifications to core of the middleware.

When receiving an unseen event, the middleware also enqueues it in timestamp
order and calls the application callback \code{cb\_sim} at the appropriate
time.
The peers must keep the full queue in memory for two reasons:
    (a) receiving a late event may reorder the queue and require a time
        travel, and
    (b) peers that join the network need to receive all events.
%
Therefore, the middleware keeps a queue of ``packets'', which includes the
event and other metadata:

@@

struct {
    int i;              // next event to apply
    int n;              // number of events
    p2p_pak buf[MAX];   // queue of packets
} PAKS;
\end{verbatim}
}

@@

The function receives the application callbacks as arguments \lin{1}
(Section~\ref{sec.tml.api}).
We first (and last) call \code{cb\_ini} \lin{2,41} to initialize (and finalize)
the application.
Then, we call \code{cb\_sim} \lin{3}, passing the event \code{START} to start
the application in real time
(Sections~\ref{sec.tml.api.events}~and~\ref{sec.tml.api.cb_sim}).
The main loop \lin{5--39} checks continuously for network packets \lin{6--23},
local input events \lin{25--31}, and time ticks \lin{33--38}.

Regarding network packets, if we receive a time synchronization packet
\code{SYNC} from the future \lin{7--10}, then we fast forward the application
to catch up in time (Section~\ref{sec.tml}.b).
If we receive an event that should have been applied in the past \lin{12--16},
then we travel back and forth (Section~\ref{sec.tml}.a).
Otherwise, we just apply the event in real time \lin{18--22}
(Section~\ref{sec.tml.api.cb_sim}).
We detail the function \code{p2p\_travel} in
Section~\ref{sec.tml.middleware.travel}.
%
Regarding local inputs, we call \code{cb\_evt} \lin{26--30} to signal if the
event should be broadcast, also transforming it to an application event
(Section~\ref{sec.tml.api.cb_evt}).
%
Regarding time ticks, we call \code{cb\_sim} and \code{cb\_ren} in real time
\lin{33--38}.
The middleware also takes periodic snapshots of the application state to
optimize time travelling (Section~\ref{sec.tml.api.state}).

@@

The function \code{p2p\_travel} time travels the application, starting at tick
\code{from}, tick by tick, until it reaches tick \code{to} \lin{2--10}.
The loop can travel in both directions, i.e., \code{from} may be higher than
\code{to}.
At each step, we need to find the tick \code{bef} with the closest past
snapshot \lin{3}, restore it \lin{4}, and than simulate the remaining steps
from the snapshot to the desired tick \lin{5--7}.
We also exhibit each step with a small \code{ms} delay \lin{8--9} so that users
experience smooth time transitions.

@@

This section provided an overview of the middleware implementation to answer,
in concrete terms, how it ensures the desired properties enumerated in the
Introduction and at the beginning of Section~\ref{sec.tml.middleware}:
    (a) peers meet event deadlines either in real time \lin{18--21} or
        travelling back and forth in time
        (\linx{12--16} / \ref{sec.tml.middleware.loop});
    (b) peers synchronize their clocks with periodic \code{SYNC} packets that
        fast forwards late peers
        (\linx{7--10} / \ref{sec.tml.middleware.loop}); and
    (c) peers can join the network at any time and still catch up in time by
        receiving the full event history
        (Section~\ref{sec.tml.middleware.events}).

@@

Regarding (a), if a peer experiences a time travel, it means that it received
an event after its deadline.
We then count all time travel occurrences during the simulation.
%
Regarding (b), we use an absolute simulation clock to track, in real time, the
execution pace of each peer.
% in comparison to the reference peer, which is the one most advanced in time.
We then count the mismatches between the clocks.
%
Regarding (c), we make peers leave and join the network periodically, possibly
creating partitions.
We then assert that they all catch up in real time within a short period.

@@

We perform simulations with 21 peers using the mixed topology of
Figure~\ref{fig.topo}, with peers organized in straight lines and cycles.
The topology has an average of 5 hops between peers (e.g., \code{4--16} are 7
hops away).
%
In order to have an absolute clock, we use a single machine to simulate all
peers, allowing us to merge their logs into a single file with a comparable
shared timeline.
%
We use \emph{NetEm}~\cite{netem} to simulate the network latency with a normal
distribution.
% (e.g., \code{tc qdisc add dev lo root netem delay 200ms 40ms distribution normal}).
%
We vary
    (i)   the network latency (\code{5--500ms} between each peer),
    (ii)  the rate of events (\code{5--200 evt/sec} in the whole network), and
    (iii) the delay multiplier (\code{1--5x} considering the latency and the
          average hops between peers).
%
Figure~\ref{fig.mult} shows the event delays considering the network latency
and chosen multiplier.
For instance, a \code{50ms} latency with a \code{2x} multiplier results in
\code{500ms} event delays, i.e., each event is triggered with this delay to
compensate the given latency and fixed average of 5 hops.
%
These variations result in $108$ combinations.
We executed each combination $3$ times for $5$ minutes each.
We also make some modifications to evaluate peer churn, which requires to
re-execute the simulation.
The experiments take around $40$ hours to complete.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{baks}
  \caption{
    \label{fig.baks}
The \code{5x}, \code{2x}, and \code{1x} delay regions each contains $36$
measures for multiple configurations of event rate (\code{5--200evt/min} and
network latency (\code{5--500ms}).
%
The color thresholds (\emph{white}, \emph{yellow}, \emph{red}, \emph{black})
are arbitrary but help to distinguish between regions of interest.
%
The experiments uses $1$ process for each peer in a single Linux machine
(\emph{i7/16GB}).
    }
\end{figure*}

We now detail the results and how we instrument the middleware implementation
to measure the properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead of time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks.
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
Figure~\ref{fig.baks} presents the results considering all combinations of
parameters (i), (ii), and (iii) above.
The colors indicate the feasibility of each tested scenario
    (\emph{white=good}, \emph{yellow=moderate}, \emph{red=poor},
    \emph{black=faulty}).
%
For instance, the three framed rectangles ($0.1$, $3.1$, and $7.2$) represent
the scenarios with $50ms$ network delay, generating $25evt/min$.
Each rectangle uses a different delay multiplier, showing a good performance
for $5x$, moderate for $2x$, and poor for $1x$.
%
\item \textbf{Real-time pace of peers:}
All peers output their current ticks continuously, allowing us to compare them
in the shared timeline.
Every time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers, which constitute time violations.
As an example, if all peers sum $1000$ of such smaller ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch at the beginning.
%
Considering all scenarios, the average is negligibly under $0.2\%$, with no
significant variance.
Therefore, we omit the resulting table akin to Figure~\ref{fig.baks}), which
would not provide further insights.
%
\item \textbf{Correctness of unstable peers:}
We tweak our simulation to make each peer remain $40s$ online and $20s$
offline in the average, including peers \code{9--11} in the middle of the
network.
This creates partitions in the network, which generate events out of sync that
require constant resynchronization between peers.
%
Then, to evaluate the network correctness under such high churn, we count how
much time the recovering peers take to catch up in real time:
When a peer becomes online, we take the current maximum tick considering all
reachable peers, and then count the time the recovering peer takes to catch up.
We also assert that they all reach the same final state eventually.
%
Considering all scenarios, the average is $1.3$ seconds of recovering time,
with no significant variance.
Therefore, we also omit the resulting table akin to Figure~\ref{fig.baks}).
%
\end{enumerate}

>>>>

As illustrated in Figure~\ref{fig.middleware}, events are broadcast between
peers with a timestamp scheduled to the future with a $\Delta$ such that all
peers are able to apply them in sync.
%
To prevent broadcast cycles, each peer keeps a vector with the highest
timestamp it received from each other peer, such that lower numbers can be
ignored when received.
Since we rely on TCP connections, it is not possible to receive out-of-order
timestamps attached to a given source peer.
When the middleware receives a higher number from a given peer, it updates its
vector position and triggers a broadcast to neighbours.
%
As mentioned in Section~\ref{sec.tml.api}, we assume that peers assign
themselves unique identifiers in a contiguous range, such that they fit in a
simple vector of timestamps.
Therefore, this is not a severe limitation, and a separate mechanism could be
used for assignments, with few modifications to the core of the middleware.

When receiving an unseen event, the middleware also enqueues it in timestamp
order and calls the application callback \code{cb\_sim} at the appropriate
time.
Peers must keep the full event queue in memory for two reasons:
    (a) receiving a late event may reorder it and require a time travel, and
    (b) peers that join the network need to receive all events.
%
Therefore, the middleware keeps a queue of ``packets'', which includes the
event and additional metadata:

@@

struct {
    int i;  // next event to apply in real time
    int n;  // number of events in the queue
    p2p_pak buf[MAX];   // queue of packets
} PAKS;
\end{verbatim}
}

@@

The function receives the application callbacks as arguments \lin{1}
(Section~\ref{sec.tml.api}).
We first (and last) call \code{cb\_ini} to initialize (and finalize) the
application \lin{2,41}.
Then, we call \code{cb\_sim} \lin{3}, passing the event \code{START} to start
the application in real time
(Sections~\ref{sec.tml.api.events}~and~\ref{sec.tml.api.cb_sim}).
The main loop \lin{5--39} checks continuously for network packets \lin{6--23},
local input events \lin{25--31}, and time ticks \lin{33--38}.

Regarding network packets, if we receive a time synchronization packet
\code{SYNC} from the future \lin{7--10}, then we fast forward the application
to catch up in time (Section~\ref{sec.tml}.b).
If we receive an event that should have been applied in the past \lin{12--16},
then we travel back and forth (Section~\ref{sec.tml}.a).
Otherwise, we just apply the event in real time \lin{18--22}
(Section~\ref{sec.tml.api.cb_sim}).
We detail the function \code{p2p\_travel} in
Section~\ref{sec.tml.middleware.travel}.
%
Regarding local inputs, we call \code{cb\_evt} to signal if the event should be
broadcast \lin{26--30}, also transforming it to an application event
(Section~\ref{sec.tml.api.cb_evt}).
%
Regarding time ticks, we call \code{cb\_sim} and \code{cb\_ren} in real time
\lin{33--38}.
The middleware also takes periodic snapshots of the application state to
optimize time travelling (Section~\ref{sec.tml.api.state}).

@@

The function \code{p2p\_travel} time travels the application, starting at tick
\code{from}, tick by tick \lin{2--10}, until it reaches tick \code{to}.
The loop can travel in both directions, i.e., \code{from} may be higher than
\code{to}.
At each step, we need to find the tick \code{bef} with the closest past
snapshot \lin{3}, restore it \lin{4}, and than simulate the remaining steps
from the snapshot to the desired tick \lin{5--7}.
We also exhibit each step with a small \code{ms} delay \lin{8--9} so that users
experience smooth time transitions.

@@

This section provided an overview of the middleware implementation to answer
how it ensures the desired properties in concrete terms as follows:
    (a) peers meet event deadlines either travelling back and forth or in real
        time (\linx{12--16} or \lin{18--22} / \ref{sec.tml.middleware.loop});
    (b) peers synchronize their clocks with periodic \code{SYNC} packets that
        fast forward late peers
        (\linx{7--10} / \ref{sec.tml.middleware.loop}); and
    (c) peers can join the network at any time and still catch up in time by
        receiving the full event history
        (Section~\ref{sec.tml.middleware.events}).

@@

Regarding (a), if a peer experiences a time travel, it means that it received
an event after its deadline.
We then count all time travel occurrences in the evaluation.
%
Regarding (b), we use an absolute simulation clock to track, in real time, the
execution pace of each peer.
% in comparison to the reference peer, which is the one most advanced in time.
We then count the mismatches between the clocks in the evaluation.
%
Regarding (c), we make peers leave and join the network periodically, possibly
creating partitions.
We then assert that they all catch up in real time within a short period in the
evaluation.

@@

We perform simulations with 21 peers using the mixed topology of
Figure~\ref{fig.topo}, with peers organized in straight lines and cycles.
The topology has an average of 5 hops between peers (e.g., \code{4-16} are 7
hops away).
%
In order to have an absolute clock, we use a single machine to simulate all
peers, allowing us to merge their logs into a single file with a comparable
shared timeline.
%
We use \emph{NetEm}~\cite{netem} to simulate the network latency with a normal
distribution.
% (e.g., \code{tc qdisc add dev lo root netem delay 200ms 40ms distribution normal}).
%
We vary
    (i)   the network latency (\code{5-500ms} between each peer),
    (ii)  the rate of events (\code{5-200 evt/sec} in the whole network), and
    (iii) the delay multiplier (\code{1-5x} considering the latency and the
          average hops between peers).
%
Figure~\ref{fig.mult} shows the event delays considering the network latency
and chosen multiplier.
For instance, a \code{50ms} latency with a \code{2x} multiplier results in
\code{500ms} event delays, i.e., each event is triggered with this delay to
compensate the given latency and fixed average of 5 hops.
%
These variations result in $108$ combinations.
We executed each combination $3$ times for $5$ minutes each.
We also make some modifications to evaluate peer churn, which requires to
re-execute the simulation.
The experiments take around $40$ hours to complete.

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{baks}
  \caption{
    \label{fig.baks}
The \code{5x}, \code{2x}, and \code{1x} delay regions each contains $36$
measures for multiple configurations of event rate (\code{5-200evt/min} and
network latency (\code{5-500ms}).
%
The color thresholds (\emph{white}, \emph{yellow}, \emph{red}, \emph{black})
are arbitrary but help to distinguish between regions of interest.
%
The experiments uses $1$ process for each peer in a single Linux machine
(\emph{i7/16GB}).
    }
\end{figure*}

We now detail the results and how we instrument the middleware implementation
to measure the properties above:
%
\begin{enumerate}
%
\item \textbf{Recurrence of time travels:}
Every time a peer is ahead of time and needs to travel back, we output its id
and how much ticks it needs to travel back.
At the end, we measure the percentage of the sum of all time travel ticks over
the total simulation ticks.
As an example, if all peers sum $1000$ time travel ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
Figure~\ref{fig.baks} presents the results considering all combinations of
parameters (i), (ii), and (iii) above.
The colors indicate the feasibility of each tested scenario
    (\emph{white=good}, \emph{yellow=moderate}, \emph{red=poor},
    \emph{black=faulty}).
%
For instance, the three framed rectangles ($0.1$, $3.1$, and $7.2$) represent
the scenarios with $50ms$ network delay, generating $25evt/min$.
Each rectangle uses a different delay multiplier, showing a good performance
for $5x$, moderate for $2x$, and poor for $1x$.
%
\item \textbf{Real-time pace of peers:}
All peers output their current ticks continuously, allowing us to compare them
in the shared timeline.
Every time we see a greater tick never seen in the timeline before, we count
smaller ticks from other peers, which constitute time violations.
As an example, if all peers sum $1000$ of such smaller ticks over $100k$
accumulated simulation ticks, then the final measure is $1\%$.
%
We also start each peer with a random delay of up to $10s$ to force a
synchronization mismatch at the beginning.
%
Considering all scenarios, the average is negligibly under $0.2\%$, with no
significant variance.
Therefore, we omit the resulting table akin to Figure~\ref{fig.baks}), which
would not provide further insights.
%
\item \textbf{Correctness of unstable peers:}
We tweak our simulation to make each peer remain $40s$ online and $20s$
offline in the average, including peers \code{9-11} in the middle of the
network.
This creates partitions in the network, which generate events out of sync that
require constant resynchronization between peers.
%
Then, to evaluate the network correctness under such high churn, we count how
much time the recovering peers take to catch up in real time:
When a peer becomes online, we take the current maximum tick considering all
reachable peers, and then count the time the recovering peer takes to catch up.
We also assert that they all reach the same final state eventually.
%
Considering all scenarios, the average is $1.3$ seconds of recovering time,
with no significant variance.
Therefore, we also omit the resulting table akin to Figure~\ref{fig.baks}).
%
\end{enumerate}
